### 2020-07-03 2-Abstract.tex

\begin{abstract}
This article presents PAMELA, an annotation-based Java Modelling framework. PAMELA provides a smooth integration between model and code, and enable Java developers to handle software development both at conceptual level and at source-code level, without code transformation and/or generation, and avoiding round-tripping issues.

% TODO...

% Parler aussi de Aspect Oriented Programming, Traits programming, Run-time weaving, Semantics Add-On, MOP (Meta Object Protocol)

\end{abstract}