### 2021-01-27 1-Main.tex
Started rebuttal process

%\documentclass[preprint]{acmart}
\documentclass[preprint]{elsarticle}
\journal{Science of Computer Programming}

%\usepackage[colorinlisttodo]{todonotes}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{tikz-uml}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2https://www.overleaf.com/project/5efee0e8ad6bc10001d47854
}

\lstset{style=mystyle}

\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}

%\acmPrice{15.00}
%\acmDOI{10.1145/1122445.1122456}
%\acmISBN{978-1-4503-9999-9/18/06}

\begin{document}
\begin{frontmatter}

\title{PAMELA: an annotation-based Java Modeling Framework}

%% Group authors per affiliation:
%\author{Sylvain Guérin, Guillaume Polet, Caine Silva, Joel Champeau}
%\address{ENSTA Bretagne}

%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard}
%\address{IMT Atlantique}

%% or include affiliations in footnotes:
\author[mymainaddress]{Sylvain Guérin\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Guillaume Polet}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Caine Silva}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Joel Champeau}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mysecondaryaddress]{Jean-Christophe Bach}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Salvador Mart\'inez}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Antoine Beugnard}
%\ead{firstname.lastname@imt-atlantique.fr}

\address[mymainaddress]{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}%vérifier si c'est l'affiliation officielle
\address[mysecondaryaddress]{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}

\input{2-Abstract}

\begin{keyword}
Model Driven Engineering \sep Object Oriented Programming \sep Metaprogramming 
\end{keyword}

\end{frontmatter}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

\section{Introduction}
\label{sec:introduction}
\input{introduction.tex}

\section{Approach}
\label{sec:pamproc}
\input{pamelaProcess.tex}

\section{Software Framework and Features}
\label{sec:architecture}
\input{architecture.tex}

\section{Implementation}
\label{sec:implementation}
\input{implementation.tex}

%\section{Industrial experimentation and validation}
%\label{sec:validation}
%\input{validation.tex}

\input{6-Experimentations}
\label{sec:Experimentations}

\section{Related Work}
\label{sec:related}
\input{related.tex}

%\input{4-ApproachOverview}

%\input{5-Features}

%\input{6-Experimentations}

%\input{7-Relatedworks}

\input{8-Conclusion}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{biblio}

\end{document}
\endinput
%%
%% End of file `sample-sigchi.tex'.

% Prout