### 2021-01-27 6-Experimentations.tex

\section{PAMELA industrial use cases and experiments}
\label{sec:validation}

 The PAMELA framework has been successfully applied in a variety of complex programming and modeling scenarios and we continue to use it daily as part of our modeling toolbox. In the following, we describe three important uses cases in which PAMELA was a core component.

%\todo[inline]{General comment: I would change the title of the section to something like, Pamela Use cases, Pamela success stories or something of the like. The idea here is to show that pamela is a tool that is used and solve problems. In that sense, the fact that pamela powers Open flexo should be highlighted. In the same sense, the importance of the patterns and security should be reduced and limited to the fact that pamela allows it (the point being, to cite the workshop paper). I believe giving details of how patterns are implemented  here will distract the reader. technical details, if needed would be better placed in the implementation section.}

%https://pamela.openflexo.org
%https://github.com/openflexo-team/pamela

\subsection{Openflexo infrastructure}

Model Federation is an approach that provides the means to integrate multiple models conforming to different paradigms, and giving to each stakeholder a specific view adapted to its needs. Model federation approach is developed as a possible response to SIMF RFP (Semantic Information Modeling for Federation) by OMG (Object Modeling Group). This RFP (Request For Proposal) requests submissions for a standard addressing, ”federation of information across different representations, levels of abstraction, communities, organizations, viewpoints, and authorities”. Thus model federation allows the integration of heterogeneous models to develop new cross-concern viewpoints/models or to synchronize the models used for designing a system.

Openflexo\cite{OpenflexoWebSite} is a software infrastructure providing support for model federation across multiple technological spaces. Conceptualization is addressed trough the proposition of a language called FML (Flexo Modeling Language), which is executable on the platform. Openflexo infrastructure introduces connectors (also called Technology Adapters) to support various technological spaces and paradigms.

This open source initiative is now mature at the infrastructure level, and many projects and applications have been developed and powered using Openflexo infrastructure. More than 15 technology adapters have been developed, with various maturity stage regarding their industrialization (Microsoft Word, Excel and PowerPoint, EMF, OWL, Diagramming, JDBC, XML, OSLC, etc...)

The total base of code for Openflexo infrastructure represents around 900.000 lines of Java code. Regarding backend modeling, the PAMELA framework is extensively used in Openflexo core as well as in most technology adapters, with very few specific implementations for properties. The observer/observable pattern is generally used for graphical user interfaces, which also rely on PAMELA framework.

% Continuous integration ?

\subsection{Formose project}

The Formose ANR (french National Agency for Research) project (ANR-14-CE28-0009)\cite{FormoseWebSite} aimed to design a formally-grounded, model-based requirements engineering (RE) method for critical complex systems, supported by an open-source environment. The main partners were: ClearSy, LACL, Institut Mines-Telecom, OpenFlexo, and THALES.

The main results of the project are a requirements modeling multi-views language, its associated design process and the development of an opensource platform called Formod\cite{FormodWebSite}, built using Openflexo infrastructure and PAMELA framework. The requirements modeling language is based on KAOS for goal modeling and SysML for the structural part of a system. The associated domain modeling language, used to describe system domain knowledge, extends the two ontology languages OWL and PLIB. The graphical notations are then translated into Event-B, a formal specification method supported by verification tools.

The Formose method and Formod tool have been evaluated on different case studies provided by the industrial partners of the project.

\subsection{SecurityPatterns experiment}

A significant experiment in PAMELA is the implementation of security patterns weaved on domain code \cite{silva20}.
In this context, the PAMELA framework is extended to include the notion of Pattern, i.e. a composition of multiple classes. Included to this experiment, the security pattern is specified by expected behavior defined and formalized by a pattern contract. This contract is defined by formal properties and the PAMELA framework ensures the property verification at run-time.

Related to the security pattern implementation, PAMELA enables the definition of additional security behavior to existing Java code.
Patterns are defined in PAMELA using three classes, each one representing a different conceptual level PatternFactory, PatternDefinition, PatternInstances.

To declare a Pattern on existing code, pattern elements such as Pattern Stakeholders and methods need to be annotated with provided security pattern-specific annotations. These annotations will be discovered at run-time by the PatternFactory and stored in PatternDefinition attributes.

Summarizing, implementing Patterns with PAMELA provides the ability to monitor the execution of the application code; the ability to offer extra structural and behavioral features, executed by the PAMELA interpreter; a representation of Patterns as stateful objects. Such objects can then evolve throughout run-time and compute assertions.

%\subsection{Performance issues}

% parler des problèmes de perfs

%\subsection{Syntaxic issues}

% Problème d'un truc trop verbeux > editeur graphique