### 2021-02-02 related.tex
typos

Very few approaches build interfaces between MDE artifacts and source code at
runtime. FAME \cite{kuhn2008fame} is  a library that keeps metamodels
accessible and adaptable at runtime. Basically, FAME attaches meta-information
to the objects of a running application permitting a uniform manipulation of
both models and system objects at runtime. Similarly, in
\cite{song2010applying} the authors use runtime models to maintain a causal
connection between a model and a running systems for monitoring and control
purposes. More recently, in \cite{boronat2019} the authors construct and
maintain at runtime model-based views on the data manipulated in
object-oriented code. Different to the aforementioned approaches, PAMELA does
not focus in reflecting and manipulating running systems but in: 1) extending
their behaviors; and 2) maintaining a continuous cycle of metamodeling so that
models and corresponding code remain synchronized without the need for code
generation processes. In a different approach, UMPLE
\cite{lethbridge2016merging} mixes programming and modeling by integrating UML
constructs into languages such as Java. However, they use code generation for
the runtime part of the system.