### 2021-02-17 2-Abstract.tex
abstract after Antidote

\begin{abstract}
This article presents PAMELA, an annotation-based Java modeling framework. PAMELA provides a smooth integration between model and code and enables Java developers to handle software development both at conceptual level and at source-code level, without code transformation and/or generation, avoiding round-trip-related issues.

% TODO...

% Parler aussi de Aspect Oriented Programming, Traits programming, Run-time weaving, Semantics Add-On, MOP (Meta Object Protocol)

\end{abstract}