### 2021-03-23 Todo-review2.tex
small corrections completed. Remain: improve fig2 comments, restructure section 2, rephrase 4.4

To trace what is corrected/or not

At the end : antidote...

Reviewer #1: The paper and presentation are much improved, and the software documentation improved somewhat, compared to the original submission.

There remain or are introduced however, a substantial number of typos, language errors, and other minor issues that need correcting:
- "to /avoid" on p3 (DONE)
- "detailed thereafter" -> "detailed hereafter" (It's Correct : As adverbs the difference between hereafter and thereafter is that hereafter is in time to come; in some future time or state while thereafter is after that, from then on.)
- "increasing PAMELA metamodel" -> "by _extending_ PAMELA_'s_ metamodel" (DONE)
- "with setter/getter": make both plural (DONE)
- POJO used here and elsewhere yet the acronym is introduced much later (DONE)
- p4 l11-12: reduce etc. should be reduce_s_ etc. (Keep as it is)
- odd spacing before : consistently throughout bullet list + text  (Keep as it is: I trust Latex)
- "in the section 3.2" -> "in Section 3.2" (style: named references deserve a capital and no article before) (DONE)
- p5 and many times: "PAMELA metamodel", "PAMELA model", "PAMELA framework" need "the" before them. (Done ; with Pamela -> PAMELA ; to be checked)
- "defines API": insert "an" (DONE)
- "in Java interface": insert "a" (DONE)
- "returning desired value", "taking value" etc.: insert "the" (DONE)
- "May also declares"... (DONE)
- p6 l46: @Embedded/"as embedded" meaning?
- p7 "of launched Java application" missing "the" (DONE)
- "is/are" -> "is (are)", in line with parentheses earlier in sentence (DONE)
- "Experimented programmers": I doubt it, unless you got major ethical clearance (Replaced by "Experienced PAMELA programmers")
- "Listing 2: model" -> "~: Model" consistent with Listing 1 (DONE)
- the book title is "Lord of the ring_s_" with s (DONE)
- "using classical Java ... scheme": missing "the" (To check :mis au pluriel)
- "Java programming language": insert "The" (DONE)
- l43 inconsistent spacing (DONE)
- l46 package ... package (DONE)
- p11 l36 sticking into margin (Final check to do)
- S.4.3 please add size in LoC and classes (?)
- java -> Java (DONE)
- p13 "diana component": insert "the" (DONE)
- "experimentation" -> "experiment" (DONE)
- "When Pamela has been" -> "After Pamela was" (DONE : if we consider it's over)
- l36 the quotation mark before code generation is the wrong one 'changed to ' and ' ?)
- p 14 "appertaining" -> "pertaining", "relating" (Kept as is. As verbs the difference between pertain and appertain is that pertain is to belong while appertain is to belong to or be a part of, whether by right, nature, appointment, or custom; to relate to.)

Reviewer #2: This new version is really better than the first one.

My major concern is related to section 2. 
- Would it be possible to structure it more, because currently it is several paragraphs the one bellow the others.
- figure 2 is not enough described in section 2 and even in the paper.

More over, concerning section 4.4, I appreciate this new section. However, while reading, it is not a good advertisement for Pamela. Overhead (x20) is huge. Perhaps it is not the right benchmark if it provides bad results that are not representing the right usage.

Minors:
- POJO should be explain the first time p3 not the second time. (DONE)
- p5 it is written twice that ModelEntities are encoded in a Java Interface (DONE)
- p13, it is written twice that Pamela framework is extensively used. (DONE : 1 changed to largely)