### 2022-03-10 main.tex
Updates from Overleaf

%\documentclass[preprint]{acmart}
\documentclass[preprint]{elsarticle}
\journal{Science of Computer Programming}

\usepackage{todonotes}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{tikz-uml}
\usepackage{hyperref}
\hypersetup{pdfauthor={Sylvain Guérin, Jean-Christophe Bach, Antoine Beugnard, Joel Champeau, Salvador Mart\'inez, Guillaume Pollet, Caine Silva}}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}

% DRAFT
\newcommand{\ab}[1]{{\color{blue} #1}} % draft mode
\newcommand{\jcb}[1]{{\color{red} #1}} % draft mode
\newcommand{\sg}[1]{{\color{purple} #1}} % draft mode

%\acmPrice{15.00}
%\acmDOI{10.1145/1122445.1122456}
%\acmISBN{978-1-4503-9999-9/18/06}

\begin{document}
\begin{frontmatter}

\title{PAMELA: an annotation-based Java Modeling Framework}

%% Group authors per affiliation:
%\author{Sylvain Guérin, Guillaume Polet, Caine Silva, Joel Champeau}
%\address{ENSTA Bretagne}

%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard}
%\address{IMT Atlantique}

%% or include affiliations in footnotes:
\author[mymainaddress]{Sylvain Guérin\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Guillaume Polet}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Caine Silva}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Joel Champeau}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mysecondaryaddress]{Jean-Christophe Bach}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Salvador Mart\'inez}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Fabien Dagnat}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Antoine Beugnard}
%\ead{firstname.lastname@imt-atlantique.fr}

\address[mymainaddress]{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}%vérifier si c'est l'affiliation officielle
\address[mysecondaryaddress]{IMT Atlantique, Lab-STICC, UMR 6285, Brest, France}

\input{abstract}

\begin{keyword}
Model Driven Engineering \sep Object Oriented Programming \sep Metaprogramming 
\end{keyword}

\end{frontmatter}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\clearpage
\section*{Software and Code Metadata}
\label{}

\section*{Current executable software version}
\label{}

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{(executable) Software metadata description} &  \\
\hline
S1 & Current software version & 1.5 \\
\hline
S2 & Permanent link to executables of this version  & https://maven.openflexo.org/artifactory/
libs-release/org/openflexo/pamela-core/1.5 \\
\hline
S3 & Legal Software License & Dual licencing: GPLV3 and EUPL \\
\hline
S4 & Computing platform/Operating System & Java runtime environment \\
\hline
S5 & Installation requirements \& dependencies & JRE \geq 1.6\\
\hline
S6 & If available, link to user manual - if formally published include a reference to the publication in the reference list & https://pamela.openflexo.org\\
\hline
S7 & Support email for questions & contact@openflexo.org\\
\hline
\end{tabular}
\caption{Software metadata}
\label{} 
\end{table}

\section*{Current code version}
\label{}

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} &  \\
\hline
C1 & Current code version & 1.6.1 \\
\hline
C2 & Permanent link to code/repository used of this code version & https://github.com/openflexo-team/pamela \\
\hline
C3 & Legal Code License   & Dual licencing: GPLV3 and EUPL \\
\hline
C4 & Code versioning system used & git / github \\
\hline
C5 & Software code languages, tools, and services used & Java \\
\hline
C6 & Compilation requirements, operating environments \& dependencies & JDK \geq 1.6\\
\hline
C7 & If available Link to developer documentation/manual & https://pamela.openflexo.org \\
\hline
C8 & Support email for questions & contact@openflexo.org\\
\hline
\end{tabular}
\caption{Code metadata}
\label{} 
\end{table}

\section{Introduction}
\label{sec:introduction}
\input{introduction.tex}

\section{Approach and Development Process}
\label{sec:pamproc}
\input{pamelaProcess.tex}

\section{Software Framework and Features}
\label{sec:architecture}
\input{architecture.tex}

\section{Implementation}
\label{sec:implementation}
\input{implementation.tex}

\input{experimentations}
\label{sec:Experimentations}

\section{Related Work}
\label{sec:related}
\input{related.tex}

\input{conclusion}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{biblio}

\end{document}
\endinput
%%
%% End of file `sample-sigchi.tex'.

% Test